### 2017-04-24 revision/figures/deform12.tex

%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{array}
\usepackage[pdftex]{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amssymb}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
$t$ & Dimensionally split & Multi-dimensional\tabularnewline
1 & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_1_T}\tabularnewline
2 & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_2_T}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/HilarysGraphics/deformingAdvectionCyclic_legends_T}}\tabularnewline
\end{tabular*}
\end{document}