### 2017-05-25 revision/makeFigs/deform_dx.tex

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular}{cc}
$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
\end{tabular}
\end{document}