### 2017-05-25 revision/makeFigs/overMountains.tex

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{array}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.14\textwidth}>{\centering}p{0.42\textwidth}c}
 & Dimensionally split & Multi-dimensional\tabularnewline
$\Delta t=25\ $s

max $c$ = 0.74

horiz $c$ =0.25

$c_{d}=0.22$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall}\tabularnewline
$\Delta t=50\ $s

max $c$ = 1.48

horiz $c$ =0.5

$c_{d}=0.44$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall}\tabularnewline
$\Delta t=100\ $s

max $c$ = 2.96

horiz $c$ =1

$c_{d}=0.87$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall}\tabularnewline
$\Delta t=200\ $s

max $c$ = 5.93

horiz $c$ =2

$c_{d}=1.75$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall}\tabularnewline
$\Delta t=500\ $s

max $c$ = 14.8

horiz $c$ =5

$c_{d}=4.4$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt500CN_up_10000_Tall}\tabularnewline
$\Delta t=1000\ $s

max $c$ = 29.6

horiz $c$ =10

$c_{d}=8.7$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt1000CN_up_10000_Tall}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.85\textwidth]{../HilarysGraphics/vSlice_horizontalAdvection_orography_legends_Tall_TdiffAll}}\tabularnewline
\end{tabular*}
\end{document}