### 2017-05-25 revision/makeFigs/sbrc1.tex

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular}{cc}
Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{../YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{../YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_legends_Tdiff}}\tabularnewline
\end{tabular}
\end{document}